\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[a4paper,height=24cm,width=13cm]{geometry}
\usepackage[italian]{babel}
\usepackage{amssymb}
\usepackage{dsfont}
\usepackage{calc}
\usepackage{graphicx}
\usepackage{pstricks}
\usepackage{pst-node}
\usepackage{fourier}
\usepackage{euscript}
\usepackage{amsmath,amssymb, amsthm}

\def\lh{\textrm{lh}}
\def\phi{\varphi}
\def\P{\EuScript P}
\def\M{\EuScript M}
\def\D{\EuScript D}
\def\U{\EuScript U}
\def\V{\EuScript V}
\def\sm{\smallsetminus}
\def\niff{\nleftrightarrow}
\def\ZZ{\mathds Z}
\def\NN{\mathds N}
\def\PP{\mathds P}
\def\QQ{\mathds Q}
\def\RR{\mathds R}
\def\<{\langle}
\def\>{\rangle}
\def\E{\exists}
\def\A{\forall}
\def\0{\varnothing}
\def\imp{\rightarrow}
\def\iff{\leftrightarrow}
\def\IMP{\Rightarrow}
\def\IFF{\Leftrightarrow}
\def\range{\textrm{im}}
\def\Mod{\textrm{Mod}}
\def\Aut{\textrm{Aut}}
\def\Th{\textrm{Th}}
\def\acl{\textrm{acl}}
\def\eq{{\rm eq}}
\def\tp{\textrm{tp}}
\def\equivL{\stackrel{\smash{\scalebox{.5}{\rm L}}}{\equiv}}
\def\swedge{\mathbin{\raisebox{.2ex}{\tiny$\mathbin\wedge$}}}
\def\svee{\mathbin{\raisebox{.2ex}{\tiny$\mathbin\vee$}}}

\newcommand{\labella}[1]{{\sf\footnotesize #1}\hfill}
\renewenvironment{itemize}
  {\begin{list}{$\triangleright$}{%
   \setlength{\parskip}{0mm}
   \setlength{\topsep}{0mm}
   \setlength{\rightmargin}{0mm}
   \setlength{\listparindent}{0mm}
   \setlength{\itemindent}{0mm}
   \setlength{\labelwidth}{3ex}
   \setlength{\itemsep}{0mm}
   \setlength{\parsep}{0mm}
   \setlength{\partopsep}{0mm}
   \setlength{\labelsep}{1ex}
   \setlength{\leftmargin}{\labelwidth+\labelsep}
   \let\makelabel\labella}}{%
   \end{list}}
%\def\ssf#1{\textsf{\small #1}}
\newcounter{ex}
\newenvironment{exercise}{\bigskip\addtocounter{ex}{1}\textbf{Exercise \theex.\quad}}{}
%\pagestyle{empty}
\parindent0ex
\parskip2ex
\raggedbottom
\renewcommand{\baselinestretch}{1.5}


\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{Example sheet 3}
\rhead{Part III}
\chead{Model Theory 2018/19}
\cfoot{}
\begin{document}



\begin{exercise}
Let $p(x)\subseteq L(A)$ and let $\phi(x\,;y)\in L(A)$ be a formula that defines, when restricted to $p(\U)$, an equivalence relation with finitely many classes (a \textit{finite equivalence relation\/}). Prove that there is a finite equivalence relation definable over $A$ that coincides with $\phi(x\,;y)$ on $p(\U)$.
\end{exercise}



\begin{exercise}
Let $T$ be a consistent theory. Suppose that all completions of $T$ are of the form 
$T\cup S$ for some set $S$ of quantifier-free sentences (for example, $T=T_{\rm acf}$). Prove that, if all completion of $T$ have elimination of quantifiers, so does $T$.

Hint: prove that for every formula $\phi(x)$ there are some quantifier-free sentences $\sigma_i$ and quantifier-free formulas $\psi_i(x)$ such that

$\displaystyle\qquad\sigma_i\vdash\phi(x)\leftrightarrow\psi_i(x)$, 
$\qquad\displaystyle T\vdash\bigvee^n_{i=1}\sigma_i$,\qquad and  $\qquad\sigma_i\vdash\neg\sigma_j$ for $i\neq j$.

For a counter example consider the empty theory in the language with a single unary predicate.
\end{exercise}



\begin{exercise}
Show that the claim in the exercise above fails when the theories $S$ have arbitrary complexity. 

Hint: let $T$ be the empty theory in the language with a single unary predicate.
\end{exercise}

\begin{exercise}
Let $\phi(z)\in L(A)$ be a consistent formula. Prove that, if $a\in\acl(A,b)$ for every $b\models\phi(z)$, then $a\in\acl(A)$. Prove the same claim with a type $p(z)\subseteq L(A)$ for $\phi(z)$.
\end{exercise}


\begin{exercise}
Let $a\in\U\sm\acl A$. Prove that $\U$ is $A$-isomorphic to some $\V\preceq \U$ such that $a\notin\V$.
\end{exercise}




\begin{exercise}\label{pofu2}
Let $C$ be a finite set. Prove that if $C\cap M\neq\0$ for every model $M$ containing $A$, then $C\cap\acl(A)\neq\0$.

Hint: by induction on the cardinality of $C$. Suppose there is a $c\in C\sm\acl(A)$, then there is $\V\simeq\U$ such that $A\subseteq\V\preceq\U$ and $c\notin\V$. Apply the induction hypothesis to $C'=C\cap\V$ with $\V$ for $\U$.
\end{exercise}


\begin{exercise}\label{ex_acl_intersezioe_2_modelli}
Prove that for every $A\subseteq N$ there is an $M$ such that $\acl A\,=\,M\cap N$.

Hint: let $\bar c$ be an enumeration of $N$. Let $p(\bar x)=\tp(\bar c/A)$. Consider the type

\hfil$p(x)\ \cup\ \Big\{\neg\big[\psi(b,\bar x)\wedge\E^{\le n}y\,\psi(y,\bar x)\big]\ :\ b\in N\sm\acl(A),\ \ \psi(y,x)\in L(A),\ \ n<\omega\Big\}$
\end{exercise}

\begin{exercise}
Let $\phi(x)\in L(\U)$ and fix an arbitrary set $A$. Prove that the following are equivalent
\begin{itemize}
\item[1.] there is some model $M$ containing $A$ and such that $M\cap\phi(\U)=\0$;
\item[2.] there is no consistent formula $\psi(z_1,\dots,z_n)\in L(A)$ such that

 \hfil$\displaystyle\psi(z_1,\dots,z_n)\imp\bigvee^n_{i=1}\phi(z_i)$.
\end{itemize}
\end{exercise}


\begin{exercise}
Prove that a in  strongly minimal theory every infinite algebraically closed set is a model.
\end{exercise}

\begin{exercise}
Prove that a countable strongly minimal theory is either $\omega$-categorical or has infinitely many countable models.
\end{exercise}

\end{document} 
